### 2008-03-19 trunk/Papers/dl-gre-08/discussion.tex
drawer pic

We will now describe two experiments evaluating different aspects of
the performance of our algorithm. First, we will look at its running
time on randomly generated domain graphs of different sizes. Second,
we will compare the output of our algorithm to human-generated
descriptions.

@@

\begin{figure}
\includegraphics[width=0.3\textwidth]{drawers.jpg}
\caption{A schematic view of the filing cabinets.\todo{needs nicer pic}}\label{fig:drawers}
\end{figure}

@@

\begin{small}
\begin{tabular}{|lp{0.9\textwidth}|}\hline
\multirow{2}{*}{drawer} & \textit{human-produce description} \\
& $\mathsf{output\ of\ the\ \el\ algorithm}$\\\hline
2 & \textit{the orange drawer above the blue drawer} \\
& (a) $\exists \mathsf{above}.(\mathsf{blue}) \sqcap \mathsf{orange}$ (b) $\exists \mathsf{above}.(\exists \mathsf{below}.(\mathsf{orange}) \sqcap \mathsf{blue}) \sqcap \mathsf{orange}$ (c) $\exists \mathsf{next}.(\mathsf{blue}) \sqcap \exists \mathsf{next}.(\mathsf{pink}) \sqcap \mathsf{orange}$\\
4 & \textit{the yellow drawer on the top of the pink one} \\ 
& (a) $\exists \mathsf{above}.(\mathsf{pink}) \sqcap \mathsf{yellow}$  (b) $\exists \mathsf{above}.(\mathsf{pink}) \sqcap \mathsf{corner} \sqcap \mathsf{green}$  (c) $\exists \mathsf{above}.(\exists \mathsf{next}.(\mathsf{green}) \sqcap \mathsf{pink}) \sqcap \mathsf{green} \sqcap \mathsf{corner}$\\
5 & $\ast$ \textit{the pink drawer in the fourth column below the yellow one}\\
& (a) $\exists \mathsf{above}.(\mathsf{orange}) \sqcap \mathsf{pink}$  (b) $\exists \mathsf{below}.(\mathsf{green}) \sqcap \mathsf{pink}$  (c) $\exists \mathsf{above}.(\exists \mathsf{next}.(\mathsf{green}) \sqcap \mathsf{orange}) \sqcap \exists \mathsf{next}.(\mathsf{green}) \sqcap \mathsf{pink}$\\
%6 & \textit{the yellow drawer on top of the yellow drawer} / \textit{the yellow drawer that's above another yellow drawer} / $\ast$ \textit{the drawer after the two blue ones in horizontal sequence}\\
6 & \textit{the yellow drawer on top of the yellow drawer} (2$\times$) / $\ast$ \textit{the drawer after the two blue ones in horizontal sequence}\\
& (a) $\exists \mathsf{above}.(\mathsf{yellow}) \sqcap \mathsf{yellow}$  (b) $\exists \mathsf{below}.(\mathsf{pink}) \sqcap \mathsf{green}$  (c) $\exists \mathsf{next}.(\mathsf{blue}) \sqcap \exists \mathsf{next}.(\mathsf{pink}) \sqcap \mathsf{green}$\\
7 & \textit{the blue drawer below the orange one} / $\ast$ \textit{the blue drawer below the orange drawer in the second column}\\ 
& (a) $\exists \mathsf{next}.(\exists \mathsf{above}.(\mathsf{orange}) \sqcap \mathsf{blue}) \sqcap \exists \mathsf{above}.(\mathsf{blue}) \sqcap \mathsf{blue}$  (b) $\exists \mathsf{below}.(\mathsf{orange}) \sqcap \mathsf{blue}$  (c) $\exists \mathsf{next}.(\mathsf{blue}) \sqcap \exists \mathsf{next}.(\mathsf{green}) \sqcap \mathsf{blue}$\\
10 & \textit{the blue drawer above the pink drawer} (2$\times$)\\
& (a) $\exists \mathsf{above}.(\mathsf{pink}) \sqcap \mathsf{blue}$  (b) $\exists \mathsf{above}.(\mathsf{pink}) \sqcap \exists \mathsf{below}.(\mathsf{blue}) \sqcap \mathsf{blue}$  (c) $\exists \mathsf{next}.(\mathsf{orange}) \sqcap \exists \mathsf{next}.(\mathsf{green}) \sqcap \mathsf{blue}$\\
%11 & \textit{the yellow drawer next to the orange drawer} / \textit{the yellow one next to the orange one}\\
11 & \textit{the yellow drawer next to the orange drawer} (2$\times$)\\
& (a) $\exists \mathsf{above}.(\mathsf{orange}) \sqcap \mathsf{yellow}$  (b) $\exists \mathsf{below}.(\mathsf{green}) \sqcap \mathsf{green}$  (c) $\exists \mathsf{next}.(\mathsf{orange}) \sqcap \mathsf{green}$\\
12 & \textit{the orange drawer below the pink drawer}\\ 
& (a) $\exists \mathsf{above}.(\mathsf{pink} \sqcap \mathsf{corner}) \sqcap \mathsf{orange}$  (b) $\exists \mathsf{below}.(\mathsf{pink}) \sqcap \mathsf{orange}$  (c) $\exists \mathsf{next}.(\mathsf{green}) \sqcap \mathsf{orange}$\\
14 & $\ast$ \textit{the orange drawer below the two yellow drawers} (2$\times$)\\
& (a) $\exists \mathsf{next}.(\mathsf{pink} \sqcap \mathsf{corner}) \sqcap \exists \mathsf{next}.(\mathsf{pink}) \sqcap \mathsf{orange}$  (b) $\exists \mathsf{below}.(\mathsf{green}) \sqcap \mathsf{orange}$  (c) $\exists \mathsf{next}.(\mathsf{pink} \sqcap \mathsf{corner}) \sqcap \mathsf{orange}$\\\hline
\end{tabular}
\end{small}

\caption{Human produced descriptions and output of our \el\  algorithm
using the orderings (a) $\mathsf{corner,}$ $\mathsf{color, above,
next, below, right, left}$, (b) $\mathsf{color, corner, below, above,
left, right, next}$ and (c) $\mathsf{corner, color, below}$
$\mathsf{above, left, right, next}$. Descriptions that the \el\
algorithm cannot generate with any ordering are marked by $\ast$.}
\label{fig:example_outputs}
\end{figure*}

@@

\newcite{viethen06:_algor_for_gener_refer_expres} also tested Dale and
Haddock's \shortcite{dale91:_gener_refer_expres_invol_relat}
Relational Algorithm on the drawer domain. Surprisingly, it could not
generate \textit{any} of the descriptions produced by humans. To
evaluate our algorithm's use of relations, we tested it on a
representation of the domain using the predicates $\mathsf{color}$ and
$\mathsf{corner}$ and the roles $\mathsf{above, below, next\ to, left\ of}$,
and $\mathsf{right\ of}$. We did not include the $\mathsf{column}$ and
$\mathsf{row}$ predicates because every drawer can be uniquely
identified through those predicates. Since our algorithm has to
consider predicates before relations, this would mean that no relation
ever gets used. Like for the Incremental Algorithm, we tested all
possible orderings of predicates and roles (with the restriction that
all predicates come before all roles). 10 of the 15 human-produced
descriptions could thus be generated using one of the 240 different
orderings. Figure \ref{fig:example_outputs} shows the output of our
\el\ algorithm for three preference orderings. Together these three
orderings cover 10 of the human-produced descriptions. 
Of the five human-produced escriptions that the \el\
algorithm can not generate, two mention column information, which we
excluded from our domain, and three involve references to sets
(\textit{the two blues ones in horizontal sequence}/\textit{the two
yellow drawers}). While our algorithm cannot reproduce those
descriptions, it does  generate other, simpler descriptions for them.

>>>>

We will now describe two experiments evaluating the quality of the \el\
algorithm's output and the efficiency of both of our algorithms, and
we discuss the interface between our algorithms and realization.

%different aspects of
%the performance of our algorithm. First, we will look at its running
%time on randomly generated domain graphs of different sizes. Second,
%we will compare the output of our algorithm to human-generated
%descriptions.

@@

\begin{figure}
\begin{center}
\includegraphics[width=0.25\textwidth]{drawers.png}
\end{center}\vspace*{-3ex}
\caption{A schematic view of the filing cabinets.}\label{fig:drawers}\vspace*{-3ex}
\end{figure}

@@

\begin{footnotesize}
\begin{tabular}{|l|p{0.95\textwidth}|}\hline
\multirow{2}{*}{id} & \textit{human-produce description} \\
& $\mathsf{output\ of\ the\ \el\ algorithm}$\\\hline
\multirow{2}{*}{2} & \textit{the orange drawer above the blue drawer} \\
&  $\bm{\mathsf{orange} \sqcap \exists \mathsf{above}.\mathsf{blue}}$ / $\mathsf{orange} \sqcap \exists \mathsf{above}.(\exists \mathsf{below}.(\mathsf{orange}) \sqcap \mathsf{blue})$ / $ \mathsf{orange} \sqcap \exists \mathsf{next}.(\mathsf{blue}) \sqcap \exists \mathsf{next}.(\mathsf{pink})$\\\hline
\multirow{2}{*}{4} & \textit{the yellow drawer on the top of the pink one} \\ 
&  $\bm{\mathsf{yellow} \sqcap \exists \mathsf{above}.\mathsf{pink}}$  / $\mathsf{yellow}  \sqcap \mathsf{corner} \sqcap \exists \mathsf{above}.\mathsf{pink}$  / $\mathsf{yellow} \sqcap \mathsf{corner} \sqcap \exists \mathsf{above}.(\exists \mathsf{next}.(\mathsf{yellow}) \sqcap \mathsf{pink})$\\\hline
\multirow{2}{*}{5} & $\ast$ \textit{the pink drawer in the fourth column below the yellow one}\\
&  $\mathsf{pink} \sqcap \exists \mathsf{above}.\mathsf{orange}$  / $\mathsf{pink} \sqcap \exists \mathsf{below}.\mathsf{yellow}$  / $ \mathsf{pink} \sqcap \exists \mathsf{next}.(\mathsf{yellow}) \sqcap \exists \mathsf{above}.(\exists \mathsf{next}.(\mathsf{yellow}) \sqcap \mathsf{orange})$\\\hline
%6 & \textit{the yellow drawer on top of the yellow drawer} / \textit{the yellow drawer that's above another yellow drawer} / $\ast$ \textit{the drawer after the two blue ones in horizontal sequence}\\
\multirow{2}{*}{6} & \textit{the yellow drawer on top of the yellow drawer} (2$\times$) / $\ast$ \textit{the drawer after the two blue ones in horizontal sequence}\\
&  $\bm{\mathsf{yellow} \sqcap \exists \mathsf{above}.\mathsf{yellow}}$  / $\mathsf{yellow} \sqcap \exists \mathsf{below}.\mathsf{pink}$  / $\mathsf{yellow} \sqcap \exists \mathsf{next}.(\mathsf{blue}) \sqcap \exists \mathsf{next}.(\mathsf{pink})$\\\hline
\multirow{2}{*}{7} & \textit{the blue drawer below the orange one} / $\ast$ \textit{the blue drawer below the orange drawer in the second column}\\ 
& $\mathsf{blue} \sqcap \exists \mathsf{above}.(\mathsf{blue}) \sqcap \exists \mathsf{next}.(\exists \mathsf{above}.(\mathsf{orange}) \sqcap \mathsf{blue})$ / $\bm{\mathsf{blue} \sqcap \exists \mathsf{below}.(\mathsf{orange})}$ \textbf{/} $\mathsf{blue} \sqcap \exists \mathsf{next}.(\mathsf{blue}) \sqcap \exists \mathsf{next}.(\mathsf{yellow})$\\\hline
\multirow{2}{*}{10} & \textit{the blue drawer above the pink drawer} (2$\times$)\\
&  $\bm{\mathsf{blue} \sqcap \exists \mathsf{above}.(\mathsf{pink})}$  / $\mathsf{blue} \sqcap \exists \mathsf{above}.(\mathsf{pink}) \sqcap \exists \mathsf{below}.(\mathsf{blue})$  / $\mathsf{blue} \sqcap \exists \mathsf{next}.(\mathsf{orange}) \sqcap \exists \mathsf{next}.(\mathsf{yellow})$\\\hline
%11 & \textit{the yellow drawer next to the orange drawer} / \textit{the yellow one next to the orange one}\\
\multirow{2}{*}{11} & \textit{the yellow drawer next to the orange drawer} (2$\times$)\\
&  $\mathsf{yellow} \sqcap \exists \mathsf{above}.\mathsf{orange}$  / $\mathsf{yellow} \sqcap \exists \mathsf{below}.\mathsf{yellow}$  / $\bm{\mathsf{yellow} \sqcap \exists \mathsf{next}.\mathsf{orange}}$\\\hline
\multirow{2}{*}{12} & \textit{the orange drawer below the pink drawer}\\ 
&  $\mathsf{orange} \sqcap \exists \mathsf{above}.(\mathsf{pink} \sqcap \mathsf{corner})$  / $\bm{\mathsf{orange} \sqcap \exists \mathsf{below}.\mathsf{pink}}$  / $\mathsf{orange} \sqcap \exists \mathsf{next}.\mathsf{yellow}$\\\hline
\multirow{2}{*}{14} & $\ast$ \textit{the orange drawer below the two yellow drawers} (2$\times$)\\
&  $ \mathsf{orange} \sqcap \exists \mathsf{next}.(\mathsf{pink} \sqcap \mathsf{corner}) \sqcap \exists \mathsf{next}.(\mathsf{pink})$  / $\mathsf{orange} \sqcap \exists \mathsf{below}.\mathsf{yellow}$  / $\mathsf{orange} \sqcap \exists \mathsf{next}.(\mathsf{pink} \sqcap \mathsf{corner})$\\\hline
\end{tabular}
\end{footnotesize}

\caption{Human produced descriptions from
\cite{viethen06:_algor_for_gener_refer_expres} for nine different drawers (id)
 and output of our \el\ algorithm using three different orderings.
 Descriptions that the \el\ algorithm cannot generate with any
 ordering are marked by $\ast$. Generated descriptions that match one
 produced by humans are in bold face. 2$\times$ indicates that the
 same description was produced twice.}

\label{fig:example_outputs}
\end{figure*}

%(a) $\mathsf{corner,}$ $\mathsf{color, above,
%next, below, right, left}$, (b) $\mathsf{color, corner, below, above,
%left, right, next}$ and (c) $\mathsf{corner, color, below}$
%$\mathsf{above, left, right, next}$. Descriptions that the \el\
%algorithm cannot generate with any ordering are marked by $\ast$.}

@@

\newcite{viethen06:_algor_for_gener_refer_expres} also tested Dale and
Haddock's \shortcite{dale91:_gener_refer_expres_invol_relat}
Relational Algorithm on the drawer domain. Surprisingly, it could not
generate \textit{any} of the descriptions produced by humans. To
evaluate our algorithm's use of relations, we tested it on a
representation of the domain using the predicates $\mathsf{color}$ and
$\mathsf{corner}$ and the roles $\mathsf{above, below, next\ to, left\ of}$,
and $\mathsf{right\ of}$. We did not include the $\mathsf{column}$ and
$\mathsf{row}$ predicates because every drawer can be uniquely
identified through those predicates. Since our algorithm (for
efficiency reasons)
considers predicates before relations, this would mean that no relation
ever gets used. Like for the Incremental Algorithm, we tested all
possible orderings of predicates and roles (with the restriction that
all predicates come before all roles). 10 of the 15 human-produced
descriptions could thus be generated using one of the 240 different
orderings. Figure \ref{fig:example_outputs} shows the output of our
\el\ algorithm for three preference orderings. Together these three
orderings cover 10 of the human-produced descriptions. 
Of the five human-produced escriptions that the \el\
algorithm can not generate, two mention column information, which we
excluded from our domain, and three involve references to sets
(\textit{the two blues ones in horizontal sequence}/\textit{the two
yellow drawers}). While our algorithm cannot reproduce those
descriptions, it does  generate other, simpler descriptions for them.
